# Your First R Project {#sec-first-project}

In this chapter you'll create your first analysis project using the Horsley Lab conventions. We'll use the `/new-project` skill to set everything up automatically.

## Open Positron

Open Positron from your Applications folder. You'll see a layout similar to RStudio but more modern:

- **Left sidebar** — file explorer, git, extensions
- **Main area** — code editor
- **Bottom panel** — terminal, console, output
- **Right sidebar** — Claude Code chat panel

## Create a Project Folder

First, create a folder for your project. In Terminal:

```bash
mkdir ~/projects/my-first-project
cd ~/projects/my-first-project
```

Then in Positron, go to **File → Open Folder** and open `my-first-project`.

## Use /new-project to Set Up the Structure

In the Claude Code sidebar in Positron, type:

```
/new-project
```

Claude will ask you a few questions:

1. **Project type** — choose **Data science**
2. **Project name** — `my-first-project` (or whatever you chose)
3. **Brief description** — one sentence about what this project is for
4. **Python needed?** — No for now
5. **Layout** — Flat
6. **Cluster?** — No for now
7. **GitHub** — Personal account, Private

Claude Code will then create the full folder structure, initialize renv, create a README, CLAUDE.md, and set up git.

::: {.callout-tip}
The `/new-project` skill sets up everything according to Horsley Lab conventions automatically. You don't need to create folders manually.
:::

## Explore the Structure

After `/new-project` finishes, you should see:

```
my-first-project/
├── data/
│   ├── raw/
│   └── processed/
├── scripts/
│   └── scratch/
├── outputs/
│   ├── figures/
│   └── tables/
├── docs/
├── .claude/
│   └── CLAUDE.md
├── renv/
├── renv.lock
└── README.md
```

## Create Your First Script

In Positron, create a new file called `01_import.Rmd` in the `scripts/` folder.

Add this content:

````markdown
---
title: "01 — Import and Explore Data"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(tidyverse)
```

## Load Data

```{r load-data}
# We'll add data loading here
cat("Script is working!\n")
```
````

## Run the Script

Click the **Run** button next to any code chunk, or use `Cmd+Shift+Enter` to run the current chunk. You should see "Script is working!" in the console.

## Use /done to Wrap Up

At the end of any working session, type in the Claude Code sidebar:

```
/done
```

This will summarize what you did, update your CLAUDE.md session log, and commit your work to git.

## Next Steps

You now have a working project with the correct Horsley Lab structure. In the next chapter, you'll learn how to work with Claude Code day-to-day to help you write and understand code.
